Where Do Events Come From?

### 2018-08-15 paper.tex
(>) Revisões Sztajnberg 2.3

The language runtime
%implements the event loop proposed in Figure~\ref{fig.async}, which
alternates between synchronous, well-behaved execution, and asynchronous,
unpredictable ISRs.
%
Since the language now has full control over the event loop and only executes
from interrupt requests, it is possible for the runtime to enter in sleep mode
automatically to save energy.
%
Listing~\ref{lst.env} is a realization of this architecture:
%
The runtime defines an input queue (ln~1) in which ISRs enqueue new events
(e.g., Listing~\ref{lst.usart.code}, ln~12).
%
In the \code{main} function, we first generate the \emph{boot reaction} (ln~3),
which starts the program, spawns the ISRs, and reaches the first \code{await}
statements in the multiple trails of the program.
%
Then, the runtime enters the infinite event loop (ln~4--11) that queries the input
queue (ln~6) to awake the program (ln~7).
%
ISRs execute asynchronously, emitting input events into the queue, which will
be queried in subsequent iterations of the loop.
%
Note that if the queue is empty, the runtime enters in sleep mode to save
energy (ln~9), and will only awake on a new hardware interrupt.

@@

Our runtime supports three compile-time configurations for \code{ceu\_pm\_sleep}
(ln~9):
%
The first configuration is a \emph{nop} that simply keeps the event loop
running all the time without ever sleeping.
This configuration is useful when introducing new platforms.
%
The second configuration chooses a (inefficient) sleep mode that keeps all
peripherals active.
Although this configuration is not the most energy efficient, at least, it
requires no extra assistance from the drivers.
%
The third sleep configuration keeps a bit vector of the active drivers during runtime
and chooses the most efficient mode, querying this vector every time
\code{ceu\_pm\_sleep} is called.

@@

input high/low PIN_02;      // connected to a button
loop do
    par/or do
        await PIN_02;
    with
        var[255] byte buf;
        await Usart_RX(&buf, 10);
        // uses buf
    end
    await PIN_02;
end
\end{lstlisting}

>>>>

The language runtime
%implements the event loop proposed in Figure~\ref{fig.async}, which
alternates between synchronous, well-behaved execution, and asynchronous,
unpredictable ISRs.
%
The runtime now also has full control over the event loop and only executes
from interrupt requests.
This allows us to propose a runtime that enters in sleep mode automatically to
save energy, as depicted in Listing~\ref{lst.env}:
%
The runtime defines an input queue (ln~1) in which ISRs enqueue new events
(e.g., Listing~\ref{lst.usart.code}, ln~12).
%
In the \code{main} function, we first generate the \emph{boot reaction} (ln~3),
which starts the program, spawns the ISRs, and reaches the first \code{await}
statements in the multiple trails of the program.
%
Then, the runtime enters the infinite event loop (ln~4--11) that queries the input
queue (ln~6) to awake the program (ln~7).
%
ISRs execute asynchronously, emitting input events into the queue, which will
be queried in subsequent iterations of the loop.
%
Note that if the queue is empty, the runtime enters in sleep mode to save
energy (ln~9), and will only awake on a new hardware interrupt.

@@

Our runtime supports three compile-time configurations for \code{ceu\_pm\_sleep}
(ln~9).
%
The first configuration is a \emph{nop} that simply keeps the event loop
running all the time without ever sleeping.
This configuration is useful when introducing new platforms.
%
The second configuration chooses a (inefficient) sleep mode that keeps all
peripherals active.
Although this configuration is not the most energy efficient, at least, it
requires no extra assistance from the drivers.
%
The third sleep configuration keeps a bit vector of the active drivers during runtime
and chooses the most efficient mode, querying this vector every time
\code{ceu\_pm\_sleep} is called.

@@

input high/low PIN_02;      // connected to a button
loop do
    par/or do
        await PIN_02;               // STATE-A
    with
        var[255] byte buf;
        await Usart_RX(&buf, 10);   // STATE-A
        // uses buf
    end
    await PIN_02;                   // STATE-B
end
\end{lstlisting}
